Low Complexity Adaptive Beamformer\\ for Active Sonar Imaging

### 2015-08-13 LCAforActiveSonarImaging/lca.tex
la2pdf commit - 2015:08:13	11:09:44	(Thu13.Aug)

\IEEEPARstart{T}{he} best adaptive sonars are found in nature. Bats, for instance, use a high frequency sonar to detect, identify and track prey with amazing precision in caves amongst a myriad of other bats. In comparison, human made sonars inhibit minimal adaptability in the data collection process, but try to make up for it in the post processing of the data. Perhaps the most studied adaptive method for reconstructing a sonar image is the minimum variance distortionless response (MVDR) beamformer~\cite{Capon1969}. It computes the window that is optimal in the sense of suppressing as much energy from the noise and interference as possible.

MVDR is often able to improve image contrast and resolution compared to conventional static methods~\cite{Blomberg2013,Blomberg2012a,Dursun2009,Lo2004}. However, this comes at a cost of a computational complexity that is cubic with the number of channels, O($M^3$), while conventional beamformers are at O($M$). We have shown in previous studies that this complexity can be reduced, and the method accelerated significantly using graphics computing units (GPUs)~\cite{Buskenes2014,Asen2013}. However, in this work we propose

We suggest using a Low Complexity Adaptive (LCA) beamformer, which for each pixel applies a set of predefined windows and selects one of them based on the MV criterion. Using the LCA beamformer results in increased robustness and improved edge definition compared to the MV and DAS beamformers.

Despite its inherent potential, the MVDR beamformer has yet to see widespread adoption in the active sonar field. There may be several reasons for this. For one, the method is not inherently robust, and may suffer from a phenomenon called signal cancellation in active systems~\cite{Widrow1982}. Another reason is that in its original form, the computational complexity is cubic with the number of channels, O($M^3$), while conventional beamformers are at O($M$). This is because a spatial covariance matrix is estimated and inverted for each image pixel.

>>>>

\IEEEPARstart{T}{he} best adaptive sonars are found in nature. Bats, for instance, use a high frequency sonar to detect, identify and track prey with amazing precision in caves amongst a myriad of other bats. Human made sonars have little to no adaptibility when data is collected, but some methods have been investigated in the data post processing. Perhaps the most studied adaptive method for reconstructing a sonar image is the minimum variance distortionless response (MVDR) beamformer~\cite{Capon1969}. It computes the window that is optimal in the sense of suppressing noise as much as possible while under the restriction of  preserving the signal of interest.

MVDR is often able to improve image contrast and resolution compared to conventional static methods~\cite{Blomberg2013,Blomberg2012a,Dursun2009,Lo2004}. However, while the computational complexity of conventional beamformers are linear with the number of channels, O($M$), MVDR is at O($M^3$). This may discourage its use and in many cases make it infeasible. We have shown in previous studies that MVDR's complexity can be reduced, as well as accelerated significantly using graphics computing units (GPUs)~\cite{Buskenes2014,Asen2013}, but there may be a way to avoid it altogether.

We propose a low complexity version of MVDR called the Low Complexity Adaptive (LCA) beamformer. This method applies a set of predefined windows and selects the one that is best able to suppress the noise. Using the LCA beamformer results in increased robustness and improved edge definition compared to the MV and DAS beamformers.

There may be several reasons for this. For one, the method is not inherently robust, and may suffer from a phenomenon called signal cancellation in active systems~\cite{Widrow1982}. Another reason is that in its original form, the computational complexity is cubic with the number of channels, O($M^3$), while conventional beamformers are at O($M$). This is because a spatial covariance matrix is estimated and inverted for each image pixel.
